Pillar 1 - Sequential Pattern Mining

### 2019-01-18 Pillar 1 - Sequential Pattern Mining.Rmd

find_course <- function(data = d_transcript, code){
  c = code
  x <- data %>%
    filter(`Course Code`== c)

@@

```
#Data Exploration of Transcripts
```{r}
exploration <- d_transcript %>%
  mutate(Grade = case_when(Grade == "NG"~ "0.0",
                           T~Grade),
         Grade=as.double(Grade))
```

@@

pass_rules <- ruleInduction(frequent_pass_trajectories, confidence = 0.1, #we need to play with the parameters.
                    control    = list(verbose = TRUE))
pass_rules_df <- as(pass_rules, "data.frame") %>%
  arrange(desc(lift))
```
Checking who failed
```{r}
view_rules <- pass_rules_df %>%
  separate(rule, into=c("LHS", "RHS"), sep = "=>", remove = T) %>% #remove= F to keep columns
  filter(str_detect(RHS, "low score")) %>% #change so that it changes automatically
  arrange(desc(confidence), desc(lift))

@@

>>>>

find_course <- function(code){ #can be modified to accept other data that is not d_transcript
  c = code
  x <- d_transcript %>%
    filter(`Course Code`== c)

@@

#Data Exploration of Transcripts
```{r}
exploration <- d_transcript %>%
  mutate(Grade = case_when(Grade == "NG"~ "-1", #"0,0", #maybe we could just remove these from the analysis and treat them separately?
                           T~Grade),
         Grade = str_replace(string = Grade, pattern = ",",replace = "."),
         Grade = as.numeric(Grade))%>%
  filter(!is.na(Grade)) %>%
  filter(Grade != -1) #remove NG

summary(exploration)

student_statistics <- exploration %>%
  group_by(`Student ID Number`) %>%
  summarise(n = n(),
            Min = min(Grade), 
            Max = max(Grade), 
            Mean = mean(Grade), 
            Median = median(Grade), 
            SD = sd(Grade), 
            `Proportion Failed` = mean(Grade < 5.5))

course_statistics <- exploration %>%
  group_by(`Course Code`) %>%
  summarise(n = n(),
            Min = min(Grade), 
            Max = max(Grade), 
            Mean = mean(Grade), 
            Median = median(Grade), 
            SD = sd(Grade), 
            `Proportion Failed` = mean(Grade < 5.5)) %>%
  inner_join(d_course, by= c("Course Code" = "Code")) #to get only UCM courses for which we have statistics

#We have NA in grades. Something weird is happening. 
summary(str_detect(d_transcript$Grade, pattern="NA")) #shows NA's = 312 FALSE= 89069 WHY ARE THERE NAs?

has_na <- d_transcript %>%
  mutate(What = str_detect(Grade, pattern="NA")) %>%
  filter(is.na(What))
```
We have found some NA values for the grades! Students:
0177849- COR1003 & EMTH0001  +++++++++++++++ Master thesis should not be showing up! +++++++++++++
0177857
0222100
0270148
0313173
0317101
0317349

#Getting graphics
```{r}

@@

pass_rules <- ruleInduction(frequent_pass_trajectories, confidence = 0.1, #we need to play with the parameters.
                    control    = list(verbose = TRUE))
pass_rules_df <- as(pass_rules, "data.frame") %>%
  arrange(desc(lift))
```
Checking who failed
```{r}
view_rules <- pass_rules_df %>%
  separate(rule, into=c("LHS", "RHS"), sep = "=>", remove = T) %>% #remove = F to keep columns
  filter(str_detect(RHS, "low score")) %>% #change so that it changes automatically
  arrange(desc(confidence), desc(lift))

@@

#Unifying transaction data
```{r}
fail_threshold <- 6.5

category <-"Low_grade" #Outcome or Low_grade

d_pass_sequence_transactions <- d_course %>%
  select(Type, Code, Letters) %>%
  filter(Type == "Elective",
         ! Letters %in% c("SKI", "PRO"),
         ! Letters %in% c("SAS", "SAH", "SAC")) %>%
  select(Code) %>%
  left_join(d_transcript, by = c("Code" = "Course Code")) %>%
  mutate(Outcome = 
           case_when( Grade=="NG"~"Fail", #if we don't do this NG satisfies the condition for pass v
                      Grade > 5.5 ~"Pass",
                              T~"Fail"),
         Low_grade =
           case_when(Grade=="NG"~"low score", #if we don't do this NG satisfies the condition for pass v
                      Grade > fail_threshold ~"high score",
                              T~"low score")
         ) %>%
  mutate(Item = paste(Code, Low_grade, sep = "_") ) %>% #Change for Outcome/Low_grade
  filter(!str_detect(Period, " to ")) %>% #filter courses that spann more than one period
  mutate( Time = paste(Year_numerical, Period, sep = ""),
          Time = as.double(Time)) %>%
  select(`Student ID Number`, Item, Time) %>%
  arrange(`Student ID Number`, Time) %>% #Note that integer identifiers must be positive and that transactions must be ordered by sequenceID and eventID.
  group_by(`Student ID Number`, Time) %>%
  summarize(list_course = list(Item)) %>% #if we run the code until here maybe its possible to use sequence_transactions <- as(sequence_transaction$list_course, "transactions")
  mutate(TransactionID = paste(`Student ID Number`, Time, sep = "_"))
```
